---
title: 'Chapter 7(Eighth Edition): 7.1, 7.2, 7.4, 7.13, 7.24, 7.25'
author: "Jeremy Ling & Emmanuel Mejia"
date: "May 1, 2018"
output: pdf_document
---

```{r,  warning=FALSE, message=FALSE}
library(car)
library("gplots")
```

##7.1

\textbf{Consider the experiment described in Problem 6.1.
Analyze this experiment assuming that each replicate represents
a block of a single production shift.}

```{r}
# defining coded
coded=function(x) #a function to code variable x
{
  ifelse(x=="+", 1, -1)
}

# creating data table
factorA = rep(c("-","+","-","+","-","+","-","+"), times = 3)
factorB = rep(c("-","-","+","+","-","-","+","+"), times = 3)
factorC = rep(c("-","-","-","-","+","+","+","+"), times = 3)
Rep = rep(c("I", "II", "III"), each = 8)
yield = c(22,32,35,55,44,40,60,39,31,43,34,47,45,37,50,41,25,29,50,46,38,36,54,47)

#dataframe
cutting.speed.long = data.frame(factorA, factorB, factorC, Rep, yield)

cutting.aov = aov(yield~Rep+factorA*factorB*factorC, cutting.speed.long)
cutting.aov.og = aov(yield~factorA*factorB*factorC, cutting.speed.long)
summary(cutting.aov); summary(cutting.aov.og)

cutting.lm = lm(yield~Rep+factorA*factorB*factorC, cutting.speed.long)
cutting.lm.og = lm(yield~factorA*factorB*factorC, cutting.speed.long)
summary(cutting.lm); summary(cutting.lm.og)
```

We look at the analysis of variance we see that the block effect is insignificant because the p-value is small, but effect B, C, and AC are significant. We also make a comparison to the analysis of Variance without the blocking to see if there was a difference. In this case there is no difference at the coefficients of our effects.

##7.2

\textbf{Consider the experiment described in Problem 6.5.
Analyze this experiment assuming that each one of the four
replicates represents a block.}

```{r}
# creating data table
A <- rep(c("-","+","-","+"), times = 4)
B <- rep(c("-","-","+","+"), times = 4)
Rep <- rep(c("I","II","III","IV"), each = 4)
Vibes <- c(18.2, 27.2, 15.9, 41.0, 18.9, 24.0, 14.5, 43.9, 12.9, 22.4, 15.1, 36.3, 14.4, 22.5, 14.2, 39.9)
router.long <- data.frame(A, B, Rep, Vibes)

# defining coded
coded=function(x) #a function to code variable x
{
  ifelse(x=="+", 1, -1)
}

# coding A and B
for (j in 1:2)
  router.long[, j]=as.numeric(coded(router.long[, j]))

#######
#router.long$Block=router.long$A * router.long$B
#router.lm = lm(Vibes ~ Block + A * B, router.long)
#router.lm.og = lm(Vibes ~ A * B, router.long)
#summary(router.lm); summary(router.lm.og)

router.aov = aov(Vibes ~ Rep + A * B, router.long)
router.aov.og = aov(Vibes ~ A * B, router.long)
summary(router.aov); summary(router.aov.og)
```

After we analyze our variance we see that all main effects are significant including the Blocking effect because all p-values are small. Blocking the replicates will create a better model for our experiment. 

##7.4

\textbf{Consider the data from the first replicate of Problem
6.1. Suppose that these observations could not all be run using
the same bar stock. Set up a design to run these observations
in two blocks of four observations each with ABC confounded.
Analyze the data.}

```{r,out.width='.49\\linewidth', fig.width=3, fig.height=3}
# creating data table
A = rep(c("-","+","-","+","-","+","-","+"), times = 1)
B = rep(c("-","-","+","+","-","-","+","+"), times = 1)
C = rep(c("-","-","-","-","+","+","+","+"), times = 1)
Rep = rep(c("I"))
yield = c(22,32,35,55,44,40,60,39)

cutting.speed.short = data.frame(A, B, C, Rep, yield)

# defining coded
coded=function(x) #a function to code variable x
{
  ifelse(x=="+", 1, -1)
}
for (j in 1:3)
  cutting.speed.short[, j]=as.numeric(coded(cutting.speed.short[, j]))

cutting.speed.short$Block = cutting.speed.short$A * cutting.speed.short$B * cutting.speed.short$C

cutting.lm_7.4 = lm(yield ~ Block + A * B  * C, cutting.speed.short)#blocking
cutting.lm_7.4.og = lm(yield ~ A * B  * C, cutting.speed.short)#original
summary(cutting.lm_7.4); summary(cutting.lm_7.4.og)#comparing coefficients

#residual analysis
res=cutting.speed.short$yield-fitted(cutting.lm_7.4)
qqPlot(res)
plot(fitted(cutting.lm_7.4), res)#bad model

#Reduced Model using probability plot
qqnorm(aov(yield ~ Block + A*B*C, cutting.speed.short), full=T,label=T)
coef=cutting.lm_7.4$coefficients[-1]
variables=names(coef)
plot=qqnorm(coef)
variables[identify(plot)]# B C AC

#New model
cutting.lm_7.4_Re = lm(yield ~ Block + A + B + C + A*C, cutting.speed.short)
summary(cutting.lm_7.4_Re)

#New Model Residual Analysis
res=cutting.speed.short$yield-fitted(cutting.lm_7.4_Re)
qqPlot(res)
plot(fitted(cutting.lm_7.4_Re), res) #better model
```

We begin by Blocing our main effects and we take a look at the coefficients of our model that includes blocking and the model that excludes blocking. Notice all coefficients are similar. To check if the model that includes the blocked data is a good model we quickly do a residual analysis and realize that our residuals are trending to a straight line, we do not have a good model, we must reduce it. Using the same method from chapter six, the probability plot, we have chosen main effects B, C, and interaction effects AC to be important effects. We apply this to a new model and check the residuals, no pattern and normality is well. The model with the Block Design data and main effects B, C, and interaction effect AC is a good model. 


##7.13

\textbf{Using the data from the $2^4$ design in Problem 6.22,
construct and analyze a design in two blocks with ABCD confounded
with blocks.}

```{r,out.width='.49\\linewidth', fig.width=3, fig.height=3}
Standard.Order = c(8,10,12,9,7,15,2,6,16,13,5,14,1,3,4,11)
Run.Order = c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16)
A = Laser.Power = c(1,1,1,-1,-1,-1,1,1,1,-1,-1,1,-1,-1,1,-1)
B = Pulse.Freq = c(1,-1,1,-1,1,1,-1,-1,1,-1,-1,-1,-1,1,1,1)
C = Cell.Size = c(1,-1,-1,-1,1,1,-1,1,1,1,1,1,-1,-1,-1,-1)
D = Writing.Speed = c(-1,1,1,1,-1,1,-1,-1,1,1,-1,1,-1,-1,-1,1)
UEC = c(0.8,0.81,0.79,0.6,0.65,0.55,0.98,0.67,0.69,0.56,0.63,0.65,0.75,0.72,0.98,0.63)
error = data.frame(Standard.Order,Run.Order,A,B,C,D,UEC)

#Blocking#
error$Block=error$A * error$B * error$C * error$D
#Linear Model#
error.lm = lm(UEC ~ Block + A*B*C*D, error)#blocking
error.lm.og = lm(UEC ~ A*B*C*D, error)#regular
summary(error.lm); summary(error.lm.og)

#Residual Analysis
res=error$UEC-fitted(error.lm)
qqPlot(res)
plot(fitted(error.lm), res)#Bad Model

#Reduced Model Procedure
qqnorm(aov(UEC ~ Block + A*B*C*D, error), full=T,label=T)
coef=error.lm$coefficients[-1]
variables=names(coef)
plot=qqnorm(coef)
variables[identify(plot)]# A C D

#New Model
error.lm.new = lm(UEC ~ Block + A + C + D, error)
summary(error.lm.new)

#Residual Analysis on New Model
res=error$UEC-fitted(error.lm.new)
qqPlot(res)
plot(fitted(error.lm.new), res)#Better Model
```

We design our data into two blocks with the confounded effects ABCD. We construct two linear models one with and the other without the Block variable. To our results we notice both models share the same coefficients. We check the residual for the Block model to see if it is a good model, unfortunately it is not, we must reduce it. We use the probability plot method to pick out which main effects are important, our results were A, C, and D. Now we apply that to a new model with the Block design and check the residuals. From our residual analysis from out new model we can state that we have a good model with main effects A, C, and D in a block design.

##7.24

\textbf{Suppose that in Problem 6.1 we had confounded ABC
in replicate I, AB in replicate II, and BC in replicate III.
Calculate the factor effect estimates. Construct the analysis of
variance table.}

```{r}
# creating data table
factorA = rep(c("-","+","-","+","-","+","-","+"), times = 3)
factorB = rep(c("-","-","+","+","-","-","+","+"), times = 3)
factorC = rep(c("-","-","-","-","+","+","+","+"), times = 3)
Rep = rep(c("I", "II", "III"), each = 8)
yield = c(22,32,35,55,44,40,60,39,31,43,34,47,45,37,50,41,25,29,50,46,38,36,54,47)

cutting.speed.long = data.frame(factorA, factorB, factorC, Rep, yield)

# defining coded
coded=function(x) #a function to code variable x
{
  ifelse(x=="+", 1, -1)
}

# subsetting Rep I
cutting.speed.long1 <- cutting.speed.long[cutting.speed.long$Rep == "I",]

A <- coded(cutting.speed.long1$factorA)
B <- coded(cutting.speed.long1$factorB)
C <- coded(cutting.speed.long1$factorC)

cutting.speed.long1$Block <- ifelse(A * B * C > 0, 1, 2)
cutting.speed.long1 <- cutting.speed.long1[order(cutting.speed.long1$Block),]

# subsetting Rep II
cutting.speed.long2 <- cutting.speed.long[cutting.speed.long$Rep == "II",]

A <- coded(cutting.speed.long2$factorA)
B <- coded(cutting.speed.long2$factorB)
C <- coded(cutting.speed.long2$factorC)

cutting.speed.long2$Block <- ifelse(A * B > 0, 1, 2)
cutting.speed.long2 <- cutting.speed.long2[order(cutting.speed.long2$Block),]

# subsetting Rep III
cutting.speed.long3 <- cutting.speed.long[cutting.speed.long$Rep == "III",]

A <- coded(cutting.speed.long3$factorA)
B <- coded(cutting.speed.long3$factorB)
C <- coded(cutting.speed.long3$factorC)

cutting.speed.long3$Block <- ifelse(B * C > 0, 1, 2)
cutting.speed.long3 <- cutting.speed.long3[order(cutting.speed.long3$Block),]

# creating new data table
p.confound <- rbind(cutting.speed.long1, cutting.speed.long2, cutting.speed.long3)
p.confound$Rep <- factor(p.confound$Rep)
p.confound$Blocks <- factor(paste(p.confound$Rep, p.confound$Block,
                                  sep = "-"))

# running regression
p.confound.aov <- aov(yield ~ Blocks + factorA * factorB * factorC, p.confound)
summary(p.confound.aov)

p.confound.lm <- aov(yield ~ Blocks + coded(factorA) * coded(factorB) * coded(factorC), p.confound)
summary(p.confound.lm)
```


##7.25

\textbf{Repeat the analysis of Problem 6.1 assuming that
ABC was confounded with blocks in each replicate.}

```{r}
# creating data table
factorA = rep(c("-","+","-","+","-","+","-","+"), times = 3)
factorB = rep(c("-","-","+","+","-","-","+","+"), times = 3)
factorC = rep(c("-","-","-","-","+","+","+","+"), times = 3)
Rep = rep(c("I", "II", "III"), each = 8)
yield = c(22,32,35,55,44,40,60,39,31,43,34,47,45,37,50,41,25,29,50,46,38,36,54,47)

cutting.speed.long = data.frame(factorA, factorB, factorC, Rep, yield)

# defining coded
coded=function(x) #a function to code variable x
{
  ifelse(x=="+", 1, -1)
}
```